---
title: "week8hw"
author: "Antonella Gonzalez"
date: "March 6, 2017"
output: md_document
---
# Exploring Data Frames
```{r}
cats <- read.csv(file = "feline-data.csv")
age <- c(2,3,5,12)
cats

age <- c(4,5,8)
cats <- cbind(cats, age)
cats

newRow <- list("tortoiseshell", 3.3, TRUE, 9)
cats <- rbind(cats, newRow)

levels(cats$coat)
levels(cats$coat) <- c(levels(cats$coat), 'tortoiseshell')
cats <- rbind(cats, list("tortoiseshell", 3.3, TRUE, 9))

str(cats)
cats$coat <- as.character(cats$coat)
str(cats)

cats[-4,]
cats
na.omit(cats)
cats <- na.omit(cats)

cats <- rbind(cats, cats)
cats

rownames(cats) <- NULL
cats
```
## Challenge 1
```{r}
df <- data.frame(id = c('a', 'b', 'c'),
                 x = 1:3,
                 y = c(TRUE, TRUE, FALSE),
                 stringsAsFactors = FALSE)

df <- data.frame(first = c('Antonella'),
                 last = c('Gonzalez'),
                 lucky_number = c(8),
                 stringsAsFactors = FALSE)
df <- rbind(df, list('Marie', 'Curie', 238) )
df <- cbind(df, coffeetime = c(TRUE,TRUE))
df
```
## Challenge 2
```{r}
#download.file("https://raw.githubusercontent.com/swcarpentry/r-novice-gapminder/gh-pages/_episodes_rmd/data/gapminder-FiveYearData.csv", destfile = "data/gapminder-FiveYearData.csv")
gapminder <- read.csv(file = "data/gapminder-FiveYearData.csv")

#source(file = "scripts/load-gapminder.R")
```

# Subsetting Data
```{r}
x <- c(5.4, 6.2, 7.1, 4.8, 7.5)
names(x) <- c('a', 'b', 'c', 'd', 'e')
x
x[1]
x[4]
x[c(1, 3)]
x[1:4]
1:4
c(1, 2, 3, 4)
x[c(1,1,3)]
x[6]
x[0]
x[-2]
x[c(-1, -5)]  # or x[-c(1,5)]

```
## Challenge 1
```{r}
x[2:4]
x[-c(1,5)]
x[c("b", "c", "d")]
```
```{r}
x[-which(names(x) == "a")]
names(x) == "a"
x[-which(names(x) %in% c("a", "c"))]
```
## Challenge 2
```{r}
x <- c(5.4, 6.2, 7.1, 4.8, 7.5)
names(x) <- c('a', 'b', 'c', 'd', 'e')
print(x)

x[-which(names(x) == "g")]
# A and C 
```
```{r}
x <- 1:3
x
names(x) <- c('a', 'a', 'a')
x

names(x) == c('a', 'c')

x[c(TRUE, TRUE, FALSE, FALSE)]
```
## Challenge 3
```{r}
x <- c(5.4, 6.2, 7.1, 4.8, 7.5)
names(x) <- c('a', 'b', 'c', 'd', 'e')
print(x)

#Write a subsetting command to return the values 4<x<7
x_subset <- x[x<7 & x>4]
print(x_subset)

```
Factor subsetting
```{r}
f <- factor(c("a", "a", "b", "c", "c", "d"))
f[f == "a"]
f[f %in% c("b", "c")]
f[1:3]
f[-3]

#Matrix subsetting
set.seed(1)
m <- matrix(rnorm(6*4), ncol=4, nrow=6)
m[3:4, c(3,1)]
m[, c(3,4)]
m[3,]
m[3, , drop=FALSE]
matrix(1:6, nrow=2, ncol=3)
matrix(1:6, nrow=2, ncol=3, byrow=TRUE)
```
## Challenge 4
```{r}
m <- matrix(1:18, nrow=3, ncol=6)
print(m)
#D. m[2,c(4,5)]
```
List subsetting
```{r}
xlist <- list(a = "Software Carpentry", b = 1:10, data = head(iris))
xlist[1]
xlist[1:2]
xlist[[1]]
xlist[["a"]]
xlist$data
```
## Challenge 5
```{r}
xlist <- list(a = "Software Carpentry", b = 1:10, data = head(iris))
xlist$b[2]

```
## Challenge 6
```{r}
#mod <- aov(pop ~ lifeExp, data=gapminder)
#attributes(mod) ## `df.residual` is one of the names of `mod`
#mod$df.residual
```

###data frames
```
head(gapminder[3])
head(gapminder[["lifeExp"]])
head(gapminder$year)
gapminder[1:3,]
gapminder[3,]
```
## Challenge 7
```
# gapminder[gapminder$year = 1957,]
gapminder[gapminder$year == 1957,]

# gapminder[,-1:4]
gapminder[,-c(1:4)]

# gapminder[gapminder$lifeExp > 80]
gapminder[gapminder$lifeExp > 80,]

# gapminder[1, 4, 5]
gapminder[1, c(4, 5)]

# gapminder[gapminder$year == 2002 | 2007,]
 gapminder[gapminder$year == 2002 | gapminder$year == 2007,]
 gapminder[gapminder$year %in% c(2002, 2007),]
``` 
## Challenge 8
Why does gapminder[1:20] return an error? How does it differ from gapminder[1:20, ]?
  -gapminder is a data.frame so needs to be subsetted on two dimensions. gapminder[1:20, ] subsets the data to give the first 20 rows and all columns.
  
Create a new data.frame called gapminder_small that only contains rows 1 through 9 and 19 through 23. You can do this in one or two steps.
  gapminder_small <- gapminder[c(1:9, 19:23),]


# Creating Publication Quality Graphics
```{r}
library("ggplot2")
gapminder <- read.csv(file = "gapminder.csv")
ggplot(data = gapminder, aes(x = gdpPercap, y = lifeExp)) +
  geom_point()

ggplot(data = gapminder, aes(x = gdpPercap, y = lifeExp))
```
## Challenge 1
```{r}
ggplot(data = gapminder, aes(x = year, y = lifeExp)) + geom_point()
```

## Challenge 2
```{r}
ggplot(data = gapminder, aes(x = year, y = lifeExp)) + geom_point(aes(color=continent))
```

Layers
```{r}
ggplot(data = gapminder, aes(x=year, y=lifeExp, by=country, color=continent)) +
  geom_line()

ggplot(data = gapminder, aes(x=year, y=lifeExp, by=country, color=continent)) +
  geom_line() + geom_point()

ggplot(data = gapminder, aes(x=year, y=lifeExp, by=country)) +
  geom_line(aes(color=continent)) + geom_point()
```

## Challenge 3
```{r}
ggplot(data = gapminder, aes(x=year, y=lifeExp, by=country)) +
 geom_point() + geom_line(aes(color=continent))
#switching the order of point and line switched which was on top 
```
Transformations and stats
```{r}
ggplot(data = gapminder, aes(x = gdpPercap, y = lifeExp, color=continent)) +
  geom_point()

#fix for outliers
ggplot(data = gapminder, aes(x = gdpPercap, y = lifeExp)) +
  geom_point(alpha = 0.5) + scale_x_log10()

ggplot(data = gapminder, aes(x = gdpPercap, y = lifeExp)) +
  geom_point() + scale_x_log10() + geom_smooth(method="lm")

ggplot(data = gapminder, aes(x = gdpPercap, y = lifeExp)) +
  geom_point() + scale_x_log10() + geom_smooth(method="lm", size=1.5)

```
## Challenge 4a and 4b
```{r}
ggplot(data = gapminder, aes(x = gdpPercap, y = lifeExp)) +
  geom_point(aes(color=continent, size=1)) + scale_x_log10() + geom_smooth(method="lm", size=1.5)

ggplot(data = gapminder, aes(x = gdpPercap, y = lifeExp)) +
  geom_point(aes(color=continent), pch=4) + scale_x_log10() + geom_smooth(method="lm", size=1.5)

```
multi-panel figures
```{r}
starts.with <- substr(gapminder$country, start = 1, stop = 1)
az.countries <- gapminder[starts.with %in% c("A", "Z"), ]
ggplot(data = az.countries, aes(x = year, y = lifeExp, color=continent)) +
  geom_line() + facet_wrap( ~ country)

ggplot(data = az.countries, aes(x = year, y = lifeExp, color=continent)) +
  geom_line() + facet_wrap( ~ country) +
  xlab("Year") + ylab("Life expectancy") + ggtitle("Figure 1") +
  scale_colour_discrete(name="Continent") +
  theme(axis.text.x=element_blank(), axis.ticks.x=element_blank())
```
## Challenge 5
```{r}

ggplot(data = gapminder, aes(x = gdpPercap, fill=continent)) +
 geom_density(alpha=0.6) + facet_wrap( ~ year) + scale_x_log10()
```






